### 2017-05-11 Simulator/article.tex
la2pdf commit - 2017:05:11	12:34:36	(to.11.mai)

\IEEEPARstart{O}{ne} of the greatest visions in science is that of artificial intelligence. If we could make machines not only do what we say, but also able to adapt their behaviour to handle the unexpected, the possibilities are perhaps only limited by our imagination. An apparent area of use is devices designed to sense and act without aid, such as autonomous underwater vehicles (AUVs). When an AUV with a sonar imaging system encounter an interesting object, it would be very beneficial if adapted its behavior to examine the object further. This is challenging as it would need to learn in what ways the interesting objects differ from the background, i.e. it relies on previous knowledge. It would also need to process this information is near real-time to act in an efficient manner.

@@

@@

The detection process is well described in the field of automatic target recognition (ATR). The classical approach involves isolating an object of interest, extracting parameters from it and feeding those into a classifier. However, it is hard to find parameters that describe the objects uniquely, in particular due to the ambiguities arising from interpreting a 3D world from its 2D projection. An alternative approach is to use template matching, where the image segment is compared to a set of template images covering the relevant object classes, and assigned to the class of the template yielding the the best fit. The challenge here is to come up with accurate template images covering the actual configurations of object and seabed, especially if a static template library is used, which will lead to a computational complexity that is exponential with the number of model parameters. This limits its use to only a few parameters being coarsely sampled, ultimately yielding inaccurate results~\cite{Midelfart2010}.

We have developed a synthetic aperture sonar (SAS) simulator that avoids this problem by running sufficiently fast to create templates adapted to the actual scene as part of the classification process. This is achieved with the aid of the massive computing power in graphics processing units (GPUs) and optimized software libraries for scene rendering. The simulator loads a 3D model of the seafloor and object class into OpenGL and positions the camera at the sonar location. The scene is highlighted using a light tube that extends to infinity along the AUV's axis of probagation. We assume rough, isotropic surfaces reflecting sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Blake1993,Bell1995}, where the intensity of the backscattered sound solely depends on the incidence angle of the transmitted signal onto the model surface. The OpenGL rendering produces an optical 2D intensity image and depth map that reveals the distance from each pixel to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.

The alternative of using of a static, predetermined template library has a computational complexity that is exponential with the number of parameters. This limits its use to only a few parameters being coarsely sampled, ultimately yielding inaccurate results~\cite{Midelfart2010}.

@@

\subsection{Setting up the scene}

@@

@@

@@

\begin{figure*}[!tbp]\centering%
\graphicsAI[width=\linewidth]{gfx/fig_images_sonar_tagged.svg}\label{fig_image_sonar}%
\caption{SAS images of 2 cylinders and a torpedo located in the Jesus Bay, Norway. Most of the 1000+ detected objects this run are similar to these.}
\end{figure*}

\begin{figure*}[!tbp]\centering%
\graphicsAI[width=\linewidth]{gfx/fig_images_simulator.pdf}\label{fig_image_simulation}%
\caption{SAS images of 2 cylinders and a torpedo located in the Jesus Bay, Norway. Most of the 1000+ detected objects this run are similar to these.}
\end{figure*}

@@

We present data from three different HUGIN runs from the Jesus Bay in Norway. The first two are taken a couple days apart in January 2009, the last in June 2012. There are roughly 3000 detections in each, most representing by cylinder shaped objects such as barrels, pipe mines and torpedoes. Therefore, the standard template library were created with SIGMAS+ using a basic  cylinder model with a set of geometrical properties commonly found in the scene\todo{Should be more precise. How many templates? How finely were the parameters sampled?}. For a fair comparison, FFISim was restricted to the same cylinder model as well.

In \Fig{fig_image_sonar} we display SAS images of three typical objects in the 2012 run. All are cylindric; the length two first being estimated to roughly 3\,m and the last 5\,m, likely representing cylinder mines and a torpedo, respectively.

@@

The classification performance is quantified with receiver operating characteristics (ROC) curves from all runs in \Fig{fig_roc_curves}. As expected, the 2009 runs proves a greater challenge than that from 2012, with 5 and 3 false positive detections at FPR=20\% in the January 6. and 8. run, respectively. This can partly be explained by incorrect estimation of orientation for a few objects in close proximity to fish trawler tracks, and partly by poor quality in highlight, shadow, or both. However, for FPR<20\% the adaptive techniques do much better than the static template library. For e.g. a mine counter measures (MCM) operation, this is the only region of interest because only a few objects out of the 3000 detections will be visually inspected up close.

>>>>

% \IEEEPARstart{W}{hen} 
% Autonomous underwater vehicles (AUVs). When an AUV with a sonar imaging system encounter an interesting object, it would be very beneficial if adapted its behavior to examine the object further. This is challenging as it would need to learn in what ways the interesting objects differ from the background, i.e. it relies on previous knowledge. It would also need to process this information is near real-time to act in an efficient manner.

@@

\IEEEPARstart{A}{utonomous} underwater vehicles (AUVs) are becoming increasingly important in human underwater endeavors. When equipped with a synthetic aperture sonar (SAS), they can e.g. be used to map the seafloor, survey underwater biology, help localize wrecks, inspect oil pipes, fields and electricity lines, or hunt for mines and biohazard waste. Such independent systems have much to gain from being able to adapt to the scenario at hand. For instance, with a well designed automatic target recognition (ATR) system on board they can adapt their mission plan to collect more information on particularly interesting objects, or fine-tune the sonar and navigation system to improve imaging performance.

@@

In conventional ATR an object of interest is first isolated, its properties attempted estimated, and a classifier used to determine what object class these properties belong to. The challenge is finding unique and intrinsic object properties, such as its type, shape and size, and to do so independently of extrinsic properties such as its alignment, burial depth, degeneration level or view angle. While SAS provide images with state of the art resolution and dynamic range, they still only represent an ambiguous acoustical 2D projection of the 3D object, littered with a highly coherent noise floor.

An alternative approach is to use template matching, where the isolated image segment is compared to a set of template images covering the relevant object classes, and assigned to the class of the template with the best fit. Since the templates are computed from 3D models, this reduces the issue of image ambiguity. However, a challenge is finding a set of templates that represent the actual configurations of object and seafloor, Another issue is that the size of precomputed library of templates grows exponentially with the number of parameters needed. This limits its use to only a few parameters being coarsely sampled, ultimately yielding inaccurate results~\cite{Midelfart2010}.

% We avoid the problems of conventional template matching by computing the templates on-the-fly as part of the classification process, thereby not needing a precomputed set of templates at all. 
%  that avoids this problem by running sufficiently fast to create templates adapted to the actual scene as part of the classification process.
% Classification process

We tackle the problems of conventional template matching by performing:
%
\begin{itemize}
\item \emph{Adaptive template matching}; where a template is selected based on object and seafloor parameters estimated from the SAS image~\cite{Midelfart2010}, and
\item \emph{Real-time template simulation}; where a template is created tailored to the estimated parameters, improving its accuracy and avoiding the need of a precomputed template library.
\end{itemize}
%

The methods presented herein will focus on the template simulator, but some details will also be given on adaptive template matching for completeness. More details on this technique can be found in~\cite{Midelfart2010}.

 This is achieved with the aid of the massive computing power in graphics processing units (GPUs) and optimized software libraries for scene rendering. The simulator loads a 3D model of the seafloor and object class into OpenGL and positions the camera at the sonar location. The scene is highlighted using a light tube that extends to infinity along the AUV's axis of probagation. We assume rough, isotropic surfaces reflecting sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Blake1993,Bell1995}, where the intensity of the backscattered sound solely depends on the incidence angle of the transmitted signal onto the model surface. The OpenGL rendering produces an optical 2D intensity image and depth map that reveals the distance from each pixel to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.

@@

Models are not ROCCAN (???)

\subsection{Loading the scene}

@@

1. Segmentation
2. Parameter estimation
   Available: Range, altitude, resolution => Input to both standard and adaptive. In standard the best fit out of a predefined template database is selected, then resampled to correct resolution.
   Additional for adaptive, parameters that modify object model: Target aspect angle, seafloor slope, object length and height.

   Target aspect angle computed with Radon Transform, seafloor under object modeled as plane with a slope estimated from bathymetric data, length estimated from longest part of highlight region, and height estimated from the shape of the shadow. The burial depth is n

3.

   Details: \ref{Midelfart2010}.

Let $S(x,y)$ be the isolated part of the

Matching the isolated image segment with the template take the form of

\begin{align}
c_{h,i} = \underset{x',y'}{\text{argmax}} \sum\limits_{\forall x}\sum\limits_{\forall y} \big| H(x,y) - T(x-x',y-y') \big|
\end{align}

\begin{align}
c_i = \frac{c_{h,i} + c_{s,i}}{2}
\end{align}

And the final template used is the one with the greatest correlation coefficient

\begin{align}
\underset{T_i(x,y)}{\text{argmax}} C_i\sum\limits_{\forall x}\sum\limits_{\forall y}  S(x,y) T(x-x',y-y')
\end{align}

Then assign object to class $i$.

@@

We tag the template matchers as either semi- or fully adaptive. The former refers to using the best matching SIGMAS+ generated template out of a precomputed set of templates, while the latter refers to the proposed method of using FFISim to generate templates on the-the-fly. \todo{Should be more precise. How many templates? Parameter span? How finely were the parameters sampled?}

@@

\begin{figure*}[!tbp]\centering%
\graphicsAI[width=\linewidth]{gfx/fig_images_sonar_simulator_tagged.svg}\label{fig_images_sonar_simulation}%
\caption{Selected objects. Top row show SAS images of 2 cylinders and a torpedo located in the Jesus Bay, Norway. Center row show the best match from a conventional template matcher, and bottom row show the adaptive template matcher. The contour of the SAS object is indicated with red lines. Note that the adaptive technique produce templates with a better fit to the highlight, but that this does not necessarily lead to a better match against the shadow.}
\end{figure*}

@@

We present data from three different HUGIN runs from the Jesus Bay in Norway. The first two are taken a couple days apart in January 2009, containing roughly 2700 and 3200 detections, respectively. The last is from June 2012 and contains roughly 650 detections. The detections represent mostly cylinder shaped objects such as barrels, pipe mines and torpedoes. Therefore, all templates were based on a basic cylinder model.

The two template matchers can be visually inspected in \Fig{fig_images_sonar_simulator_tagged}.

The classification performance is quantified with receiver operating characteristics (ROC) curves from all runs in \Fig{fig_roc_curves}. The 2009 runs proves a greater challenge than that from 2012, with 5 and 3 false positive detections at FPR=20\% in the January 6. and 8. run, respectively. This came as no surprise due to the 2009 runs containing fish trawler tracks and having poorer highlight and shadow quality than the run from 2012. However, for FPR<20\% the adaptive techniques outperform the static template library. For e.g. a mine counter measures (MCM) operation, this is the only region of interest because only a few objects out of the 3000 detections will be visually inspected up close.

In \Fig{fig_images_sonar_simulator_tagged} we compare the SAS image to the templates generated by the conventional and adaptive template matcher. Three typical objects from the 2012 were chosen, all being cylindric. The length two first being estimated to roughly 3\,m and the last 5\,m, likely representing cylinder mines and a torpedo, respectively.

Selected objects. Top row show SAS images of 2 cylinders and a torpedo located in the Jesus Bay, Norway. Center row show the best match from a conventional template matcher, and bottom row show the adaptive template matcher. The contour of the SAS object is indicated with red lines. Note that the adaptive technique produce templates with a better fit to the highlight, but that this does not necessarily lead to a better match against the shadow.

@@

### 2017-05-11 ThesisJo/introduction.tex
la2pdf commit - 2017:05:11	12:34:36	(to.11.mai)

@@

>>>>

\IEEEPARstart{O}{ne} of the greatest visions in science is that of artificial intelligence. If we could make machines not only do what we say, but also able to adapt their behaviour to handle the unexpected, the possibilities are perhaps only limited by our imagination.

@@

JIB
Smarter, better, faster - On the future of machine intelligence in active sonar systems

Preface
  Way too complex a topic to be covered in fully detail. Perhaps impossible?
  Interesting enough to be covered, even if just partially.
  Complexity issue - picture from essay

Introduction
   A marvel of science, the world, and beyond
      Intelligence / conscience. What is it?
      HPC bridging learning and big data
   Machine intelligence / adaptivity / Deep learning
      Means for intelligence?
      Adapts to data
      Proven powerful
      Hard to understand, hard to control
      Data hungry
      Fascinating perspectives:
      - Local dependencies work on e.g. images. Due to lack of data?
      - First layer similar to template matching (convolitional network)
      - If we can't understand it, perhaps we can parent it? Society's norms a giant assemble of prior knowledge.
      - Deep neural nets interconnected, learning from each other?
      - The brain adapts its wiring to solve tasks, does this carry over?
      - Neuman networks - recursive
   Big data
      Fundament of intelligence?
      Amount of data recorded and stored immense. Sensors everywhere.
      Phones, internet searches, smart watches, cars, surveillance cams,   
   High performance computing
      Engine of intelligence?
      Combining deep learning with big data
      Improving at a rapid pace

Background
   Acoustic waves
      Wave equation
   Image reconstruction
      Conventional beamforming
      Adaptive beamforming
         LCA
         MVDR
   Image modeling
      Images based on prior knowledge
      Can be used as input to deep learning

Summary of publications
   - Red line: HPC, adaptivity,
   Journal:
      2017 - JOE - Simulator

      2017 - JOE - Low-Complexity Adaptive Sonar Imaging
      2015 - JOE - An optimized GPU implementation of the MVDR beamformer for active sonar imaging
      2014 - TUFFC - Implementing capon beamforming on a GPU for real-time cardiac ultrasound imaging

   Related work:
      2014 - UAM2014 - A GPU sonar simulator for automatic target recognition
      2013 - ICA2013 - Adapting the minimum variance beamformer to a graphics processing unit for active sonar imaging systems
      2012 - ECUA2012 - GPU-Based Adaptive Beamforming for Active Sonar Imaging
      2012 - IEEE-US - Implementing Capon Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging
      2011 - UAM2011 - A low complexity adaptive beamformer for active sonar imaging

Discussion and future work
   Summary and discussion
      Simulator:
      - The way to train and constrain deep learning algorithms?
      MVDR:
      - Analytical optimization - full adaptivity.
      - Heavy on computations
      LCA:
      - Optimization by trial and error - discrete solution space.
      - Works almost as well as MVDR in most cases
      - Fast!
      - Well suited for hardware

      Lessions from deep learning:
      -

   Future work
